Developing a Visual Studio Code Language Support Extension for the Snail Programming Language

### 2023-06-15 paper.tex
feat: initial complete draft of section 4

A language server is a tool used to provide language-specific editing features such as advanced autocomplete, go-to definition, or automatic error checking \cite{Microsoft_2022a}. Many IDEs today utilize language servers. A development tool will launch a separate language server process to run in the background while a developer writes code. The development tool will then send a \emph{request} to the language server for advanced diagnostics, and the language server will \emph{respond} with the information requested by the development tool. The development tool will then display the diagnostics to the developer. This process is displayed in figure Figure \ref{fig:language-server}.

Many language servers are implemented in the language they assist with. For example, a typical Python language server would be implemented in Python. Fruther, a single language server might have to interact with a number of different development tools. For example, a Python language server might have to interact with VS Code, PyCharm, and Spyder. Without a standard framework for communication between development tools and language servers, development tools and language servers must implement specific functionality for each pairing. This problem is displayed in Figure \ref{fig:no-lsp-theory}.

@@

The Langauge Server Protocol (LSP) defines a standard framework for development tool and language server communication \cite{Microsoft_2022a}. This way, a development tool simple sends and recieves information according to the LSP standard, and is able to communicate with any language-specific language server this way. Similarly, a langauge server will be able to communicate with any development tool by communicating via the LSP standard.

@@

TODO describe the difference between events, requests, and notifications

@@

@@

\begin{wrapfigure}{C}{0.3\textwidth}
    \includegraphics[width=0.3\textwidth]{lsp-sequence.png}
    \caption{Test A}
    \label{fig:lsp-sequence}
\end{wrapfigure}

>>>>

A language server is a tool used to provide language-specific editing features such as advanced autocomplete, go-to definition, or automatic error checking \cite{Microsoft_2022a}. Many IDEs today utilize language servers. A development tool will launch a separate language server process to run in the background while a developer writes code. The development tool will then send a \emph{request} to the language server for advanced diagnostics, and the language server will \emph{respond} with the information requested by the development tool. The development tool will then display the diagnostics to the developer. This process is displayed in figure \ref{fig:language-server}.

Many language servers are implemented in the language they assist with. For example, a typical Python language server would be implemented in Python. Further, a single language server might have to interact with a number of different development tools. For example, a Python language server might have to interact with VS Code, PyCharm, and Spyder. Without a standard framework for communication between development tools and language servers, development tools and language servers must implement specific functionality for each pairing. This problem is displayed in Figure \ref{fig:no-lsp-theory}.

@@

The Langauge Server Protocol (LSP) defines a standard framework for development tool and language server communication \cite{Microsoft_2022a}. This way, a development tool simply sends and recieves information according to the LSP standard, and is able to communicate with any language-specific language server this way. Similarly, a langauge server will be able to communicate with any development tool by communicating via the LSP standard.

@@

VS Code's LSP relies on JSON-RPC requests and responses \cite{Microsoft_2023i}. A request is a message that a development tool sends to a language server. This message contains an identifier, method to be executed, and paramaters to be provided to that method. In order to follow the LSP, a language server \textbf{must} send the corresponding response message upon recieving any requests. This response message contains the request identifier, the result of the executed method, and an error object if the request fails. For example, a development tool might send the \lstinline{Initialize} request to a language server, and a language server must respond with the \lstinline{InitializeResponse} before the development tool is allowed to send any further requests. Language servers are also allowed to send requests to development tools, and development tools are still required to respond with the corresponding response.

VS Code's LSP also contains notifications. A notification is a message that does not require a response. A development tool can send a notification to a language server and the language server can process the notification as it sees fit. Often, the language server will send information back to the development tool (in the form of another notification), but it is important to note that the language server is not required to respond.

@@

\begin{wrapfigure}{C}{0.3\textwidth}
    \includegraphics[width=0.3\textwidth]{lsp-sequence.png}
    \caption{Modelling the interactions between a developer, a VS Code client, and snail language server while a developer opens and modifies a snail file.}
    \label{fig:lsp-sequence}
\end{wrapfigure}

@@
